# Comparison to other methods

```{r, echo=FALSE}
knitr::opts_chunk$set(error=FALSE, message=FALSE)
```

## Setting up the comparisons

This is a evaluation of how `spacker` does on small positive integer data compared to general purpose compression methods.
We will compare both speed and the compression ratio.

```{r}
COMPARE <- function(x) {
    x0 <- spacker:::encode_short(x)

    gz.time <- system.time(gz.out <- memCompress(x0, type='gzip'))[["elapsed"]]
    bz.time <- system.time(bz.out <- memCompress(x0, type="bzip2"))[["elapsed"]]
    xz.time <- system.time(xz.out <- memCompress(x0, type="xz"))[["elapsed"]]
    spkd.time <- system.time(spkd.out <- spacker::spack(x, method="doubling"))[["elapsed"]]
    spkm.time <- system.time(spkm.out <- spacker::spack(x, method="multiplier"))[["elapsed"]]

    gz.time.out <- system.time(memDecompress(gz.out, type='gzip'))[["elapsed"]]
    bz.time.out <- system.time(memDecompress(bz.out, type="bzip2"))[["elapsed"]]
    xz.time.out <- system.time(memDecompress(xz.out, type="xz"))[["elapsed"]]
    spkd.time.out <- system.time(spacker:::doubling_unpacker(spkd.out, length(x)))[["elapsed"]]
    spkm.time.out <- NA# system.time(spacker:::multiplier_unpacker(spkm.out, length(x)))[["elapsed"]]

    packing <- c(
        SpackDoubling = length(spkd.out),
        SpackMultiplier = length(spkm.out),
        Gzip = length(gz.out),
        Bzip2 = length(bz.out),
        Xz = length(xz.out)
    )

    times <- c(
        spkd.time, 
        spkm.time,
        gz.time, 
        bz.time, 
        xz.time
    )

    otimes <- c(
        spkd.time.out, 
        spkd.time.out,
        gz.time.out, 
        bz.time.out, 
        xz.time.out
    )

    data.frame(Ratio=packing/length(x)/4, Time=times, Unpack=otimes)
}
```

## Simulating some data

Generating some small positive integer data (a.k.a., mostly 1's).
`spacker`'s doubling method is in the same ballpark as the other methods in terms of compression, though it is much simpler and faster.

```{r}
y <- rpois(1e6, 1) + 1
table(y)
COMPARE(y)

y <- rpois(1e6, 0.5) + 1
table(y)
COMPARE(y)

y <- rpois(1e6, 0.1) + 1
table(y)
COMPARE(y)
```

It deteriorates as soon as the integers get a bit larger, though.
`spacker`'s assumption of monotonic decreasing frequencies is violated.

```{r}
y <- rpois(1e6, 10) + 1
COMPARE(y)

# Restoring that assumption slightly.
tab <- table(y)
z <- sort(tab, decreasing=TRUE)
remapped <- match(y, as.integer(names(z)))
COMPARE(remapped)
```

## Playing with real data

The real use-case for `spacker` is in transmitting highly sparse single-cell data, where most of the values are 1.
Here, we'll have a look at a few datasets.

```{r}
library(scRNAseq)
sce <- BacherTCellData()
counts <- assay(sce)@x
COMPARE(counts[1:1e6])

deltas <- abs(diff(assay(sce)@i))
COMPARE(deltas[1:1e6])
```

```{r, echo=FALSE}
rm(sce)
gc()
```

```{r}
sce <- ZilionisLungData()
counts <- assay(sce)@x
COMPARE(counts[1:1e6])

deltas <- abs(diff(assay(sce)@i))
COMPARE(deltas[1:1e6])
```

```{r, echo=FALSE}
rm(sce)
gc()
```

```{r}
sce <- ZeiselNervousData()
counts <- assay(sce)@x
COMPARE(counts[1:1e6])

deltas <- abs(diff(assay(sce)@i))
COMPARE(deltas[1:1e6])
```

```{r, echo=FALSE}
rm(sce)
gc()
```

## Session information

```{r}
sessionInfo()
```
